Thesis Title

### 2018-01-05 cuthesis_phd.tex
Added some 'notes' on the HDFS paper.. to be fixed

\documentclass{report}
\usepackage{cuthesis}
\usepackage{hyperref}
\usepackage{xcolor}
\author{Valerie Hayot-Sasson}
\title{Thesis Title}
\titleOfPhDAuthor{Ms.}          % or Ms., Mrs., Miss, etc. (only for PhD's)
\PhD                            % Masters by default
\degree{Computer Science}
% \dept{Computer Science}       % default is Comp.Sci.
\cosupervisor                   % if you also have a co-supervisor

@@

\chapter{Parallelization frameworks}
\section{Databases}
\section{Pipelining engines}
\chapter{MapReduce}
Summary of ~\cite{mapred}

@@

In order to ensure fault tolerance, the master regularly pings the worker nodes. Should a worker node not respond after a certain delay, the master labels the worker as failed. The \textit{map} tasks that were completed by the worker are all returned to their original state (\textit{idle}) as \textit{map} task data is stored locally. Both the completed and in-progress worker's \textit{map} tasks are subsequently reassigned to other workers. Only the failed worker's \textit{in-progress} \textit{reduce} tasks need to be reassigned.

@@

When nearing completion of the MapReduce program, it is possible for  ``stragglers" (\textit{in-progress} tasks running on nodes with below average performance) to delay completion. As a measure to counteract this, the master node assigns the same task to idle workers. The task is then marked as completed as soon as either the primary or the backups complete the task. This strategy has been found to significantly improve performance when used with very large datasets. \note{44\% speedup with their sort program}.

@@

\section{Hadoop}
\cite{hadoop}

@@

\chapter{Pipelining engines used in neuroscience}
\cite{nipype}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Body of Thesis goes here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\documentclass{report}
\usepackage{cuthesis}
\usepackage{hyperref}
\usepackage{xcolor}
\author{Valerie Hayot-Sasson}
\title{Thesis Title}
\titleOfPhDAuthor{}          % or Ms., Mrs., Miss, etc. (only for PhD's)
\PhD                            % Masters by default
\degree{Computer Science}
% \dept{Computer Science}       % default is Comp.Sci.
\cosupervisor                   % if you also have a co-supervisor

@@

\chapter{Introduction}
intro here
\chapter{Parallelization frameworks}
\section{Databases}
\section{Pipelining engines}
\chapter{MapReduce}
Summary of ~\cite{mapred}

@@

In order to ensure fault tolerance, the master regularly pings the worker nodes. Should a worker 
node not respond after a certain delay, the master labels the worker as failed. The \textit{map} 
tasks that were completed by the worker are all returned to their original state (\textit{idle}) as 
\textit{map} task data is stored locally. Both the completed and in-progress worker's 
\textit{map} tasks are subsequently reassigned to other workers. Only the failed worker's \textit{in-
progress} \textit{reduce} tasks need to be reassigned.

@@

When nearing completion of the MapReduce program, it is possible for  ``stragglers" (\textit{in-
progress} tasks running on nodes with below average performance) to delay completion. As a 
measure to counteract this, the master node assigns the same task to idle workers. The task is 
then marked as completed as soon as either the primary or the backups complete the task. 
This strategy has been found to significantly improve performance when used with very large 
datasets. \note{44\% speedup with their sort program}.

@@

\note{Brief hadoop section??}

\section{Hadoop Distributed File System}
Summary of \cite{hadoop}

The Hadoop Distributed File System (HDFS) is the filesystem component of Hadoop. The 
metadata in HDFS is stored in on a dedicated server known as the NameNode, whereas 
application data is distributed across many servers referred to as DataNode. To ensure fault 
tolerance, HDFS replicates the data across multiple DataNodes (default replication factor: 
3). Not only does this ensure that the data is not lost in the event of node failure, but it also increases data transfer bandwidth, as the data can be accessed from multiple nodes, and thus, there are more opportunities for computations to be performed nearest to where the data is located.

The NameNode contains the namespace tree, a hierarchy of directories and their files, as well as the mapping of the split file blocks to DataNodes. The entire namespace is stored in memory. Information on the directories and files, such as modification and access times, namespace and disk quotes, are stored within \textit{inodes} on the NameNode. The name system's metadata, \textit{inode} and file block mapping, is known as the \textit{image}. Persistent record of the image that is stored on the local file system are known as \textit{checkpoint}. Locations of block replicas are not stored in the checkpoint as they may change over time. The \textit{journal} is a log of the modifications made to the image. Both the checkpoint and the journal may be copied across servers for increased durability. The journal is played back during the NameNode restart in order to restore the cluster.

When a client requests to read data, it must first contact the NameNode. The NameNode provides it with the replica locations, and the client reads from the DataNode located closest to it. When the client requests to write a file, it contacts the NameNode which selects the DataNodes that will host the replicas. The client subsequently writes the data directly to the DataNodes in a pipeline fashion (The data gets propagated through each DataNode by being transferred from the nearest DataNode to it).

There is only one NameNode assigned to each cluster, however, each cluster can have multiple clients and execute multiple tasks concurrently.

A block replica stored on a DataNode consists of two files stored on the local file system: the metadata and the data.

The DataNode connects to the NameNode during startup and performs handshake to for DataNode namespace ID and software version verification. Should either namespace ID or software version not match those of the NameNode, the DataNode shuts down.

To ensure integrity of the system, a \textit{namespace ID} is assigned to all nodes during formatting of the namespace. DataNodes with different namespace IDs to the NameNode will not be permitted to join the cluster. Should the DataNode have not been assigned a namespace ID, it will be permitted to join the cluster. The cluster's namespace ID will then be assigned to that DataNode.

Following the handshake, the DataNode register with the NameNode using their storage ID. The storage ID is a unique ID generated after initial registration to the NameNode and that is persistently stored on the DataNode. It permits identification of the DataNode in the event of an IP address or port change.

inodes, stored on namenode, represent file and directories. inodes record attributes such as permissions, modifications and access times, namespace and disk quotas.

namenode keeps track of namespace tree and mapping of file blocks to datanodes.

The \textit{image} stores the \textit{inode} data and the list of blocks belonging to each file.

The \textit{checkpoint} is a persistent record of the \textit{image} stored in the local filesystem.

The \textit{journal} is the modification log stored on the local filesystem.

@@

\section{Schedulers}
\chapter{Pipelining engines used in neuroscience}
\cite{nipype}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Body of Thesis goes here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
